Aldebaran~b's temperate past uncovered in planet search data

### 2018-04-05 paper/aldebaran.tex
new calculations

With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
Using the same \citet{2017Rod} MESA models as were used to compute the stellar parameters, interpolating with the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 3.24~Gyr from $0.82\, \pm\, 0.27$ to $1.25^{+0.52}_{-0.40}$ times that of the Earth today (Figure~\ref{insolation}). After $\sim 3.2$ Gyr, the high mass tail of the simulations have evolved off the main sequence and dominate estimates for the insolation, while masses at the low end permit a substantially longer duration on the main sequence ($\sim Gyr$). Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.

>>>>

With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
Using the same \citet{2017Rod} MESA models as were used to compute the stellar parameters, interpolating with the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 3.24~Gyr from $0.82\, \pm\, 0.27$ to $1.25^{+0.52}_{-0.40}$ times that of the Earth today (Figure~\ref{insolation}). After $\sim 3.2$ Gyr, the high mass tail of the simulations have evolved off the main sequence and dominate estimates for the insolation, while masses at the low end permit a substantially longer duration on the main sequence ($\sim$~Gyr). Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf, i.e. planets orbiting the brown dwarf and not the primary or wider binary) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
